---
title: "Nowcasting revisions using the generalized Kishor-Koenig model"
output: rmarkdown::html_vignette
bibliography: references.bib
biblio-style: apalike
link-citations: true
vignette: >
  %\VignetteIndexEntry{nowcasting-revisions-kk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Having some evidence that the revisions are predictable, we can now move to the next step: **nowcasting**. Nowcasting refers to the process of estimating the current state of the economy using timely indicators. In the context of revisions, nowcasting involves predicting the final (or efficient) value of an economic variable before all revisions are available. This vignette demonstrates how to implement a nowcasting model using the generalized Kishor-Koenig (KK) framework [@kishorVAREstimationForecasting2012].

## The Generalized Kishor-Koenig Model

The KK model extends the traditional Vector Autoregression (VAR) framework to account for data revisions explicitly. Unlike conventional models that treat data as final and accurate, the KK model recognizes that initial releases are subject to revisions, leading to biased and inefficient forecasts. By incorporating the revision process into a state-space framework, the KK model provides a reliable approach to nowcasting economic variables. Moreover, it nests several nowcasting models, such as the classical textbook measurement error model or the @howreyUsePreliminaryData1978 model.

The procedure is based on the assumption that there exists an efficient estimate $y_t^e$ of the final release that becomes available $e$ periods after the initial release $y_t^0$. We show in vignette [Efficient Release Identification](efficient-release.html) how the number of the efficient release can be empirically tested. We follow @strohsalDataRevisionsGerman2020 and assume that $e$th revision data follows an autoregressive model of order one (AR(1)). The KK model is represented by the following equations:

**1. The State Equation**
\[ z_t = F z_{t-1} + \nu_t \]

**2. The Observation Equation**
\[y_t = (I - G) F y_{t-1} + G z_t + \epsilon_t\]

where 
\[ z'_t = \begin{bmatrix} y_{t-e}^e, y_{t-e +1}^e, ...,  y_t^e \end{bmatrix}' \]
\[ y'_t = \begin{bmatrix} y_{t-e}^e, y_{t-e +1}^{e-1}, ...,  y_t^0 \end{bmatrix}' \]
\[ \nu'_t = \begin{bmatrix} 0, 0, ...,  \nu_{0,t} \end{bmatrix}' \]
\[ \epsilon'_t = \begin{bmatrix} 0, \epsilon_{e-1,t}, ...,  \epsilon_{1,t}, \epsilon_{0,t} \end{bmatrix}' \]
\[
F = \begin{bmatrix} 0 & 1 & 0 & \dots & 0 \\
                      0 & 0 & 1 & \dots & 0 \\
                      \vdots & \vdots & \vdots & \ddots & \vdots \\
                      0 & 0 & 0 & \dots & 1 \\
                      0 & 0 & 0 & \dots & F_0 \end{bmatrix}
\]
\[
G = \begin{bmatrix} 1 & 0 & \dots & 0 \\
                      G_{e-1,e} & G_{e-1,e-1} & \dots & G_{e-1,0} \\
                      G_{e-2,e} & G_{e-2,e-1} & \dots & G_{e-2,0} \\
                      \vdots & \vdots & \ddots & \vdots \\
                      G_{0,e} & G_{0,e-1} & \dots & G_{0,0} \end{bmatrix}
\]
\( G \) is a gain matrix capturing the weight placed on new information, \( F \) determines the AR coefficient, and \( \nu_t \) and \( \epsilon_t \) are error vectors. The state-equation and observation-equation error vectors are assumed to be uncorrelated with one another at all leads and lags, and to be serially uncorrelated.

Because \( \nu_t \) and \( \epsilon_t \) are serially uncorrelated, and uncorrelated with one another at all leads and lags, estimation of the model equations is unproblematic apart from the crossequation restrictions on \( F \).
Therefore, it is recommended to estimate this system using seemingly unrelated regression (SUR). The SUR estimation is based on the [systemfit::nlsystemfit()] function. The parameter space of the KK model increases exponentially with the number of the efficient release. This leads to long computation times and potential convergence issues. As a fast, but potentially biased alternative, the package also allows to estimate the models parameters equation by equation using ordinary least squares (OLS). 


### Howrey and Classical Measurement Error Model
It is easy to see that the KK model nests the classical measurement error model and the @howreyUsePreliminaryData1978 model. 
measurement error model treats published data ($y_t$) as a noisy measure of the true economic variable ($z_t$), where $y_t = z_t + \eta_t$. It addresses uncertainty in interpreting recent data by optimally updating forecasts, considering both potential economic shocks and measurement errors. This model is obtained by setting \( G = I \).

As an extension to the classical measurement error mode, @howreyUsePreliminaryData1978 proposed a model that allows the revisions to be autocorrelated. Setting the entries $G_{e-i,0} = 0$ for $i = 1,...,e-1$ and $G_{0,0} = 1$ results in the Howrey model.


### State Space Representation
Then, armed with estimates of F and K, one can apply the Kalman filter to the following state-space model (notation following @durbinTimeSeriesAnalysis2012)

**1. The Observation Equation**
\[ y_t = Z \alpha_{t} + \varepsilon_t \]

where \( \varepsilon_t \dist N(0,H)\). @durbinTimeSeriesAnalysis2012 assumed that both error terms are iid and orthogonal to one another. However, we write the model in the form where \( \varepsilon_t \), and therefore also \( H \) are equal to zero and may be omitted, so this equation becomes

\[ y_t = Z \alpha_{t} \]

**2. The State Equation**
\[ \alpha_{t+1} = T \alpha_{t} + R \eta_{t} \]

where
\[ \alpha_t = \begin{bmatrix} z_t \\ y_t - z_t  \end{bmatrix} \]
\[ Z = \begin{bmatrix} I & I \end{bmatrix} \]
\[ T = \begin{bmatrix} F & 0 \\ 0 & (I-G)F \end{bmatrix} \]
\[ R = I \]
\[ \eta_t = \begin{bmatrix} \nu_t \\ \epsilon_t - (I-G) \nu_t \end{bmatrix} \]
\[ \eta_t \sim N(0, Q) \]
 
The analyst sees truth and the revisions. Both truth and revisions are included in the state vector. Importantly, when calculating the gain matrix, the model requires that one allow for the possibility that the off-diagonal matrix elements of \( Q \equiv E[\eta_t \eta'_t] \) are nonzero. In particular, our model implies that 
 \[ Q = \begin{bmatrix} \nu_t \nu'_t & -\nu_t \nu'_t (I-G)'\\ - (I-G) \nu_t \nu'_t & \varepsilon_t \varepsilon'_t + (I-G) \nu_t \nu'_t (I-G)' \end{bmatrix} \]



### Forecasting
Kalman filter forecasts are generated by projecting the estimated state vector forward in time using the model's estimated transition equation $\langle T \rangle$. Given the Kalman filter's state estimate $\langle \alpha_t \rangle_T$ at time $T$, forecasts for future periods $T+i$ are calculated as:

$$\langle \alpha_{T+i} \rangle_T = \langle T \rangle_T^i \langle \alpha_T \rangle_T$$


### Nowcasting using the `reviser` package
As a demonstration, we nowcast the Euro Area GDP using the KK model. We use the `reviser::gdp` dataset, which contains the Euro Area GDP data. As a first step we test which release (among the first 15) is efficient. As final release, we use the value published 4 years after the initial value, allowing for a 4-year revision period. It turns out that $e=2$ meaning that the third release is an efficient estimate of the final release.


```{r warning = FALSE, message=FALSE}
library(reviser)
library(magrittr)
library(dplyr)
library(lubridate)

gdp <- reviser::gdp %>%
  tsbox::ts_pc() %>%
    dplyr::filter(id == "EA",
      time >= min(pub_date),
      time <= as.Date("2020-01-01")
  ) %>%
  tidyr::drop_na()

df <- get_nth_release(gdp, n = 0:14)

final_release <- get_nth_release(gdp, n = 15)

efficient_release <- get_first_efficient_release(
    df,
    final_release
  )

data <- efficient_release$data
e <- efficient_release$e

summary(efficient_release)
```


We then use the `kk_nowcast()` function to nowcast the Euro Area GDP using the KK model. The function requires a `df` with columns `time`, `value`, `release#` and `final`. The function `get_first_efficient_release()` returns a dataset in this format. 


```{r warning = FALSE, message=FALSE}
nowcast <- kk_nowcast(
    df = data,
    e = e,
    model = "KK",
    method = "MLE"
  )

# Estimated parameters
nowcast$params

# Filtered states
filt_states <- nowcast$states
tail(filt_states)
```

Now we can plot the forecast.

```{r warning = FALSE, message=FALSE, include=FALSE, eval=FALSE}
plot(nowcast)
```


### References
